The effect of climate forcing on numerical simulations of the Cordilleran ice sheet at the Last Glacial Maximum

### 2014-03-31 reviews/disc-ac1.tex
Add minor corretions from Arjen to responses

In fact, no. In Figs.~8 and~9 of the discussion paper, temperature climatologies are compared without lapse-rate correction. Due to differences in resolution of surface topography between climatological datasets used, this caused appearance of numerous topographic artefacts on paper Fig.~9. Following your comment, we produced updated figures that include a lapse-rate correction of 6\,{$^\circ$}C\,km$^{-1}$, as applied in the simulations.

The implementation of lapse-rate correction in Fig.~9 (temperature difference maps, Fig.~AC1) results in smoother temperature difference maps, where effective temperature discrepancies are no longer overshadowed by local topographic anomalies (Fig.~AC2). Because temperature differences in this new map are much reduced, we adjusted colour scaling to emphasize smaller values.

Also, the new Fig.~8 (temperature density maps) now uses bilinearly-interpolated temperature maps, rather than the original data (as presented in Fig.~2--4). This results in smoother density plots than those presented in the discussion paper (Fig.~AC3). In a second step, we apply a lapse-rate correction to project all reanalysis data onto the WorldClim (higher resolution) topography (Fig.~AC4). It now becomes clear that much of the discrepancies observed were due to lapse-rate effects.

For consistency, Fig.~10 (precipitation density maps) was also updated using bilinearly-interpolated data (this comment's Fig.~AC5). Note that for all density maps, colour scales were changed to allow for a higher level of detail.

While these new figures present an altered version of climate forcing data used in our study, they more closely reflect how this data is read in by the ice sheet model, which seems appropriate for the discussion section. Moreover, the new maps reveal more clearly the strengths and weakness of different datasets, and support more strongly our interpretation of the ``hybrid'' climate forcing experiments presented in Sect.~5.2. We will definitely include these new figures in the revised manuscript. Thank you very much for this constructive comment.

@@

We will correct this.

@@

To avoid confusion, we will use negative values in the revised manuscript.

@@

A short discussion of these effects is given in Sect.~5.3, in the context of comparison of simulation results to the geomorphological last glacial maximum ice margin. However you raise here a valid concern regarding potential ice sheet evolution to more similar states when accounting for precipitation changes, not considered in the current version of our manuscript. In the East more particularly, where in most of our simulations, the ice margin does not attain a steady-state configuration, its position after 10\,kyr is largely determined by its rate of advance during the simulation length, which in turn depends on the amount of precipitation received. We agree that precipitation corrections in this region could slow down the ice margin advance, resulting in more similar configuration after 10\,kyr. We will include a discussion of this effect in Sect.~5.5.

>>>>

In fact, no. In Figs.~8 and~9 of the discussion paper, temperature climatologies are compared without lapse-rate correction. Due to differences in resolution of surface topography between climatological datasets used, this caused appearance of numerous topographic artefacts in Fig.~9. Following your comment, we produced updated figures that include a lapse-rate correction of 6\,{$^\circ$}C\,km$^{-1}$, as applied in the simulations.

The implementation of a lapse-rate correction in Fig.~9 (temperature difference maps, Fig.~AC1) results in smoother temperature difference maps, where effective temperature discrepancies are no longer overshadowed by local topographic anomalies (Fig.~AC2). Because temperature differences in this new map are much reduced, we adjusted colour scaling to emphasize smaller values.

Also, the new Fig.~8 (temperature density maps) now uses bilinearly-interpolated temperature maps, rather than the original data (as presented in Figs.~2--4). This results in smoother density plots than those presented in the discussion paper (Fig.~AC3). In a second step, we apply a lapse-rate correction to project all reanalysis data onto the WorldClim (higher resolution) topography (Fig.~AC4). It now becomes clear that much of the discrepancies observed were due to lapse-rate effects.

For consistency, Fig.~10 (precipitation density maps) was also updated using bilinearly-interpolated data (Fig.~AC5). Note that for all density maps, colour scales were changed to allow for a higher level of detail.

These new figures present an altered version of climate forcing data used in our study, and they now more closely reflect how this data is read in by the ice sheet model, which seems appropriate for the discussion section. Moreover, the new maps reveal more clearly the strengths and weakness of different datasets, and support more strongly our interpretation of the ``hybrid'' climate forcing experiments presented in Sect.~5.2. These new figures are now included in the revised manuscript. Thank you very much for this constructive comment.

@@

We have corrected this.

@@

To avoid confusion, we have adopted negative temperature offset values in the revised manuscript and figures where appropriate.

@@

A short discussion of these effects is given in Sect.~5.3, in the context of comparison of simulation results to the field evidence of the last glacial maximum ice margin. However you raise here a valid concern regarding potential ice sheet evolution to more similar states when accounting for precipitation changes, not considered in the current version of our manuscript. More particularly, in the east, where in most of our simulations, the ice margin does not attain a steady-state configuration, its position after 10\,kyr is largely determined by its rate of advance during the simulation length, which in turn depends on the amount of precipitation received. We agree that precipitation corrections in this region could slow down the ice margin advance, resulting in more similar configuration after 10\,kyr. We have included a discussion of this effect in Sect.~5.5.
